R for Clinical Study Reports and Submission

### 2021-08-20 project-folder.Rmd
update project information

# Project Folder

In a late-stage clinical trial, the number of A&R deliverables can easily be in the hundreds. 
For an organization, it is possible to have multiple ongoing clinical trials in a clincial program.

@@

## Reproducibility

@@

If the R version mismatch, an error message is displayed.

>>>>

# Overview

In a late-stage clinical trial, the number of A&R deliverables can easily be in the hundreds. 
For an organization, it is also common to have multiple ongoing clinical trials in a clinical program.

In this part, we would like to put you as an A&R leader for a clinical project.

In Chapter \@ref(folder), we will discuss how to organize source code, documents, 
deliverables in an A&R clinical project. 
We recommend to use the R package folder structure.

In Chapter \@ref(manage), we will discuss a process or system development lifecycle 
to manage A&R of a clinical project. 
We recommend to follow an agile management approach to define, develop, validate and deliver work.

# Project Folder {#folder}

@@

## Reproducibility {#reproduce}

@@

If the R version mismatch, an error message is displayed as below.

### 2021-08-20 project-management.Rmd
update project information

# Project Management

- ToDo

A clearly defined software development lifecycle (SDLC) for the programming of A&R deliverables is crucial to manage all
the outputs and to facilitate the development of the deliverables.
A typical SDLC for the A&R deliverables is comprised of 4 stages: define, develop, validate and operations.
The define stage is the most critical stage in the lifecycle as the requirements for all A&R deliverables are
gathered and documented. Validation plan for the deliverables is also planned and documented at this
stage. After the define stage is complete, the programming development occurs in the develop stage. The
programs completed in the develop stage are then validated according to the validation plan in the validate
stage. Test runs of the completed programs are conducted to ensure quality of the deliverables.
The development of programs and their associated validation is usually carried out in a dedicated test area
in the computing platform. After the validation stage is completed, the operations stage commences with
the promotion of validated programs to the production area in the computing platform. The promoted
programs are then executed in the production area with production data to generate the final version of the
deliverables. The deliverables are then handed over to the end customer for review. Any changes in
requirements of the deliverables at this stage based on end customer review, results in the deliverable
being processed through the 4 stages of SDLC again, starting with the changes to the specifications in the
define stage.
Several different software and programming languages have used in analyzing clinical trial data, with R
being among them. R is a popular programming language and an open-source software environment. R
has been widely used in various fields of data analytics for statistical analyses. There is a growing interest
within the pharmaceutical industry in analyzing clinical trial data using R and its associated software under
a regulatory environment. However, there is a gap in defining the appropriate SDLC to complete typical
A&R deliverables and ensure regulatory compliance. 
ToDo:

- add mkstdpkg/mkdev/mkuse related information.

>>>>

# Project Management {#manage}

For A&R deliverables in a clinical project development, 
a clearly defined process or system development lifecycle (SDLC) 
is crucial to ensure regulatory compliant.

SDLC for A&R deliverables can be defined in stages:

- planning: a planning stage to define the scope of a project.  
- development: a development stage to implement target deliverables. 
- validation: a validation stage to verify target deliverables. 
- operation: an operation stage to deliver work to stakeholders.

The define stage is an important stage in the SDLC lifecycle 
as the requirements for all A&R deliverables are gathered and documented.

Importantly, we should not consider SDLC as a linear process. 
For example, if study team identify new requirement in a development or validation stage, 
team shall return to planning stage to discuss and align the scope. 
An [agile project management](https://www.atlassian.com/agile/project-management) approach is suitable 
and recommended to an A&R clinical development project.
The goal is to embrace an iterative approach that continue improve target deliverables based on
frequent customer feedback.

There are many good tools to implement agile project management strategy, for example:

- [github project board](https://docs.github.com/en/issues/organizing-your-work-with-project-boards/managing-project-boards/about-project-boards)
- [Jira](https://www.atlassian.com/software/jira)

## Planning

In the planning stage, a project leader shall identifies all the deliverables, 
e.g. a list of table, listing, and figures (TLFs). 
For each TLFs, team shall prepare necessary specifications:

- mock-up
- validation level (e.g. independent review or double programming)
- etc.

ToDo: add a simplified mock-up and A&R grid.

project leader shall also align work assignment with team members.
The purpose is to answer the question on "who is doing what?".

ToDo: add a simplified validation tracker

Project lead shall also setup a project folder as discussed in Chapter \@ref(folder).
The project initiation can be simplified by creating an 
[RStudio project template](https://rstudio.github.io/rstudio-extensions/rstudio_project_templates.html).

To enable reproducibility, project leader shall also review the startup file 
(i.e. `.Rprofile` discussed in Section \@ref(manage)) and define:

- R version 
- Repository of R packages with a snapshot date 
- Project package library path 
- etc.

> After project initiate, modifying `.Rprofile` will be a risk of reproducibility 
> and shall be handled carefully if necessary.

## Development

After a project is initiated, the study team start to develop TLFs based on 
pre-defined mock-up tables assigned to each team member.

The analysis code and relevant description can be saved in R markdown files 
in the `vignettes` folder. 
The use of R markdown allow developer to assemble narrative text, code and its
comments in one place to simplify documentation. 
It would be helpful to define a name convention for all TLFs deliverables. 
For example, we can use `tlf_` prefix in the filename to indicate that the Rmarkdown is to deliver TLFs. 
For example, `tlf_001_ds.Rmd` is an Rmarkdown file for disposition table. 
Multiple TLFs with similar design can be included in one Rmarkdown file.

ToDo: add a file path for `esubdemo`

To validate the results, 
developer shall also save the last dataset before creating TLFs in RTF format to `.Rdata` files.

## Validation

## Operation
